Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-08-03 paper/ms.tex
update ms

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust or gas clumps in orbit at
  the Keplerian corotation radius.  Here, we report new CQVs
  discovered in TESS short-cadence data collected between July~2018
  and Sep~2022.  Our search of \nstarssearched\ K and M dwarfs with
  $T$$<$16 and $d$$<$150\,pc yielded \ngoods\ high-quality CQVs.  Most
  of these discoveries are new, and they include the brightest
  ($T$$\approx$9.5) and closest ($d$$\approx$20\,pc) examples of this
  class of object known, as well as the most massive
  ($\approx$0.6-0.8\,$M_\odot$) and the oldest ($\approx$200\,Myr).  A few
  objects are outliers among outliers; LP 12-502 for instance shows a
  recurring ``dip complex'' with a period and total duration that are
  fixed over more than 1{,}000 cycles, but which in detail contains
  anywhere from four to eight local minima per cycle, and at times
  displays distinct superposed periods simultaneously.  Broadly
  speaking, we find that none of the CQVs maintain a fixed light curve
  shape over timescales of more than a few hundred cycles, and we
  revisit the arguments for why transient corotating material is the
  most viable explanation.  We expect that our sample will facilitate
  future efforts aimed at connecting these objects to the broader
  contexts of star, disk, and perhaps even exoplanet evolution. 
  %For instance, if the transiting material is dust, then the rate of
  %dip evolution implies that of order an asteroid belt's worth of mass
  %passes through these structures over their lifetime of
  %$\approx$$10^8$ years.
\end{abstract}

@@

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified phenomenologically
using their optical light curves, which show nearly periodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle.  Most are
pre-main-sequence M dwarfs without near-infrared excesses, ages of
$\approx$5-150 million years (Myr), and rotation periods of at most
two days; they are observed to be $\approx$1-3\% of M dwarfs younger
than 100 Myr \citep{2016AJ....152..114R,2022AJ....163..144G}.  The
dips can be chromatic, with a reddening law plausibly consistent with
dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

@@

Given this context, it is tempting to try to interpret features of the
CQV light curves in terms of the possible presence of close-in
exoplanets, or even planetesimals.  Rocky planets this young could have
molten global magma oceans analogous to those that existed on the
Earth and Moon \citep[see][]{2022arXiv220310023L}, and thus might be
undergoing significant outgassing and atmospheric escape.  The system
would then be analogous to the Jupiter-Io plasma torus \citep[e.g.][]{1980JGR....85.1171N}, with a similar rotation
period for the primary and orbital period for the secondary,
but magnetic fields stronger by a factor of $\approx$1{,}000.
While this type of configuration is {\it a priori} plausible,
any evidence in support of such a picture currently seems
ambiguous at best.

>>>>

% 250 word max!
\begin{abstract}
  Complex quasiperiodic variables (CQVs) are low-mass
  pre-main-sequence stars with nearly periodic optical modulation.
  The modulation is likely induced by dust or gas clumps in orbit at
  the Keplerian corotation radius.  Here, we report new CQVs
  discovered in TESS short-cadence data collected between July~2018
  and Sep~2022.  Our search of \nstarssearched\ K and M dwarfs with
  $T$$<$16 and $d$$<$150\,pc yielded \ngoods\ high-quality CQVs.  Most
  of these discoveries are new, and they include the brightest
  ($T$$\approx$9.5) and closest ($d$$\approx$20\,pc) examples of this
  class of object known, as well as the most massive
  ($\approx$0.6-0.8\,$M_\odot$) and the oldest ($\approx$200\,Myr).  A few
  objects are outliers among outliers; LP 12-502 for instance shows a
  recurring ``dip complex'' with a period and total duration that are
  fixed over more than 1{,}000 cycles, but which in detail contains
  anywhere from four to eight local minima per cycle, and at times
  displays multiple periods simultaneously.  Broadly
  speaking, we find that none of the CQVs maintain a fixed light curve
  shape over timescales of more than a few hundred cycles, and we
  revisit the arguments for why transient corotating material is the
  most viable explanation.  We expect that our sample will facilitate
  future efforts aimed at connecting these objects to the broader
  contexts of star, disk, and perhaps even exoplanet evolution. 
  %For instance, if the transiting material is dust, then the rate of
  %dip evolution implies that of order an asteroid belt's worth of mass
  %passes through these structures over their lifetime of
  %$\approx$$10^8$ years.
\end{abstract}

@@

Data from K2 and TESS have yielded a new class of variable star whose
root cause is only beginning to become clear: complex quasiperiodic
variables (CQVs).  These objects are identified 
from their optical light curves, which show nearly periodic troughs
that can be either sharp or broad, often superposed on smooth
spot-like modulation
\citep{2017AJ....153..152S,2018AJ....155...63S,2019ApJ...876..127Z}.
Some CQVs show up to eight local minima (dips) per cycle.  Most are
pre-main-sequence M dwarfs without near-infrared excesses, ages of
$\approx$5-150 million years (Myr), and rotation periods of at most
two days; they are observed to comprise $\approx$1-3\% of M dwarfs younger
than 100 Myr \citep{2016AJ....152..114R,2022AJ....163..144G}.  The
dips can be chromatic, with a reddening law plausibly consistent with
dust
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
And finally, while the dip shapes can ``jump'' between different
depths and durations over less than one cycle, they more often evolve
gradually, over tens to hundreds of cycles
\citep[e.g.][]{2017AJ....153..152S,2022ApJ...925...75P,2023ApJ...945..114P}.

@@

Given this context, it is tempting to try to interpret features of the
CQV light curves in terms of the possible presence of close-in
exoplanets, or even planetesimals.  Rocky planets this young could
have molten global magma oceans analogous to those that existed on the
Earth and Moon \citep[see][]{2022arXiv220310023L}, and thus might be
undergoing significant outgassing and atmospheric escape.  The system
would then be analogous to the Jupiter-Io plasma torus
\citep[e.g.][]{1980JGR....85.1171N}, with a similar rotation period
for the primary and orbital period for the secondary, but magnetic
fields stronger by a factor of $\approx$1{,}000.  While this type of
configuration is {\it a priori} plausible, no direct evidence in
support of such a picture currently exists.
